Tight Bounds for Online Balanced Graph Partitioning

### 2020-05-08 paper/disc20clustering.tex
minor
wip

\noindent
\textbf{Learning model.}
We study a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.
We refer to this problem variant as the \emph{learning} problem.

@@

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	Note that \OPT cannot have both pairs collocated at the same time.
	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume  \OPT has one pair,
	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case $\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
	\begin{comment}
	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
	Latter,
	 we argue that \OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		let $\OPTM_1 (R_t) := 2\alpha$.
		\item
		\OPT has one of the pairs  already collocated and the other pair separated,
			  prior to and during their respective intervals.
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$
		and let $\OPTM_2 (R_t) := \lambda$.
		\item
		\OPT has one of the pairs already collocated prior to its interval,
		and it collocates the other pair during the respective interval.
		Then it pays 0 for one pair and $\alpha$ for the other and 
		let $\OPTM_3 (R_t) := \alpha$.
	\end{enumerate}
%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
	If \OPT behaves in one of these cases,
	then
	$\mathit{OPT} (R_t) \geq
		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	\OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
	Therefore in any case,
	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
	\end{comment}

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$ (separating $\{u,v\}$).
	We define times $t'' < t' < m$ similarly as in the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\mathit{OPT} (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.

>>>>

\noindent
\textbf{Learning model.}
We study a variant of Online Balanced Partitioning, where communication requests are permanent: once a pair of nodes start communicating, they continue forever.
To avoid infinite cost, each communicating pair must be collocated.
We study instances with finite cost: we assume that the communication graph can be perfectly partitioned in the clusters (no external requests are required).
In the partitioning model, $\alpha$ denoted the ratio between the cost of a single communication request and the cost of migration; in the learning model, we set $\alpha = 1$.
The algorithm's goal is to \emph{learn} the communication graph, and partition it without too many reconfigurations.

@@

	Next,
	we bound  costs incurred by \OPT for requests that trigger collocation of $\{u,v\}$ at $t \in T$.
	If $t$ is the first time that \RM collocates $\{u,v\}$ then  \OPT pays
	$\lambda$ for serving requests in $\sigma_{\set{u,v}}[0,t]$ (remotely),
	or $\alpha$ for collocating the pair and
	serving (some of) the requests with  cost zero.
	Therefore in this case,
	$\OPTM (\sigma_{\{u,v\}}(0,t]) \geq  \min{\{ \lambda,\alpha \}}$.
	%
	Otherwise,
	it is not the first collocation and
	consider times $t'$ and $t''$ as define previously,
	 and let 
	$R_t := \sigma_{\{u,w\}}(t'',t'] \cup \sigma_{\{u,v\}}(t',t] $.
	We define $R_{t'}$ for the collocation at $t'$  analogously.
	Then,
	$\OPTM(R_t) = \mathit{OPT} (\sigma_{\{u,w\}}) 
	+ \OPTM(\sigma_{\{u,v\}}) $.
	Note that \OPT cannot have both pairs collocated at the same time.
	However, it can have one of the pairs already collocated prior to its respective interval.
	Let us assume  \OPT has one pair,
	e.g.~$\{u,v\}$ collocated already prior its respective interval, i.e.~$(t',t]$,
	 and keeps it so during the interval.
	 Then it pays zero while serving $\sigma_{\{u,v\}}$.
	Hence,
	it must pay $\alpha$ for collocating the other pair, in this case $\{u,w\}$,
	or (resp., and) it pays (resp., up to) $\lambda$ for serving (resp., some of) requests in $\sigma_{\{u,w\}}$. 
	Therefore in any case,
	$\OPTM(R_t) \geq  \min{\{ \lambda,\alpha \}}$.
	\begin{comment}
	We distinguish three (non-exhaustive) cases for the cost to \OPT w.r.t.~$R_t$.
	Latter,
	 we argue that \OPT cannot save cost by deviating from these cases.
	\begin{enumerate}[label=\roman*.]
		\item
		\OPT collocates $\{u,w\}$ at  $t''$, $\{u,v\}$  at $t'$,
		and keeps them so during $(t'',t']$ and $(t',t]$, respectively.
		Therefore it serves all requests in $R_t$ with no additional cost and
		let $\OPTM_1 (R_t) := 2\alpha$.
		\item
		\OPT has one of the pairs  already collocated and the other pair separated,
			  prior to and during their respective intervals.
		Hence,
		it pays only $\lambda$ for serving all requests in $R_t$
		and let $\OPTM_2 (R_t) := \lambda$.
		\item
		\OPT has one of the pairs already collocated prior to its interval,
		and it collocates the other pair during the respective interval.
		Then it pays 0 for one pair and $\alpha$ for the other and 
		let $\OPTM_3 (R_t) := \alpha$.
	\end{enumerate}
%	We consider cases (ii) and (iii) also for the pair $\{u,w\}$, symmetrically.
	If \OPT behaves in one of these cases,
	then
	$\mathit{OPT} (R_t) \geq
		\min{\{ \OPTM_1(R_t), \OPTM_2(R_t), \OPTM_3(R_t) \}}  =
		\min{ \{ \lambda, \alpha \}}$.
	Otherwise,
	\OPT deviates from these cases,
	e.g.~by collocating a pair after paying for some of requests to this pair.
	However,
	no deviation can decrease the cost to \OPT. \mahmoud{prove this!}
	Therefore in any case,
	$\mathit{OPT} (R_t) \geq  \min{ \{ \lambda, \alpha \}}$.
	\end{comment}

	It remains to bound the cost  incurred by \OPT due to requests to $\{u,v\}$ that do not lead to its collocation until the end of the sequence at $t=m$.
	We bound the cost analogously to the case where \RM collocates $\{u,v\}$.
	If $\{u,v\}$ is not collocated in the initial matching
	and \RM never collocates it,
	then $ C^{m}_{ \{u,v\} } =| \sigma_{\{u,v\}}[1,m] |$.
	\OPT pays
	$\mathit{OPT} (\sigma_{\{u,v\}}[1,m]) 
	\geq \min{ \{ \alpha, C^{m}_{ \{u,v\} } \} }$,
	for collocating this pair or (and) paying for (resp.~some of) requests in $\sigma_{\{u,v\}}[1,m]$.
	Else,
	either $\{u,v\}$ is collocated in the initial matching
	or \RM collocates it at some point.
	Then there exists an intersecting pair $\set{u,w}$
	that is collocated by \RM at $t' < m$,
	separating $\{u,v\}$.
	We define times $t'' < t' < m$ similarly as in the former case.
	Let $R^*_{\{u,v\}} := \sigma_{\{u,w\}} (t'',t'] \cup \sigma_{\{u,v\}} (t',m]$.
	Then,
	\OPT must pay for collocating at least one pair or (and) serving requests 
	to the other pair remotely.
	Thus,
	$\mathit{OPT} (R^*_{\{u,v\}}) 
	\geq  \min{ \{ C^{m}_{ \{u,v\}}, \alpha \}}$.
